---
title: "Customer Churn"
author: "Please see the pdf for the details"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(BDgraph)
library(survival)
data(churn)
summary(churn)
```

The churn rate, also known as the rate of attrition or customer churn, is the rate at which customers stop doing business with an entity. It is most commonly expressed as the percentage of service subscribers who discontinue their subscriptions within a given time period

The churn dataset contains the length of time the customer was with the com-
pany (Account.Length), whether they churned (Churn) and a number
of other covariates for 3333 customers.

(a) Construct a Kaplan-Meier estimate of the survival function for
time-to-churn and plot the estimate of the survival function.
```{r}
 # Fit Kaplan-Meier Curve
churn$Churn=ifelse(churn$Churn=="True",1,0)

# Set up Surv() object
survdat <- Surv(time=churn[,2],event=churn[,20])

# Fit the model to the data
fit <- survfit(survdat~1,se=FALSE)
#summary(fit)
plot(fit, main= "Survival Function Estimation", xlab="Account Length", ylab="Rate")

```

(b) Estimate the median time-to-churn.
```{r}
#median
print(fit)
```

(c) Use the Cox proportional hazards method to study the impact
of the customer covariates on the time-to-churn and develop a
parsimonious model.
```{r}
# Fit the Cox PH model to the data
fit <- coxph(Surv(Account.Length,Churn)~.,data=churn)

# Examine the fit
summary(fit)
```

```{r}
# Plot the baseline survival function
library(ggplot2)
library(survminer)
survfit(fit) 
```

```{r}
obj<-step(fit)
summary(obj)
```
